---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
## Essai du Modele sans test ni validation set

```{r echo=TRUE, message=FALSE, warning=FALSE}
library("randomForestSRC")
library("survival")
library("pec")

#all_data <- read.csv(file="Onehot_wNA.csv", header=TRUE, sep="\t", na.strings="NA", dec=",",row)
all_data <- read.csv(file="Onehot_noNA_v2.csv", header=TRUE, sep=";", na.strings="NA", dec=",",row.names=1)

all_data$my_Hscore<-NULL

#all_data$my_Hscore<-NULL

n_tree_values <- list(43,45,47)
n_split_values <- list(4,5)
split_rule_values <- list("logrank") 
n <- names(all_data)
surv <- as.formula(paste("Surv(OS,STATUS)~",paste(n[!n %in% "y"], collapse = " + ")))

x<- NULL
best_n_tree <-NULL
best_n_split <- NULL
best_split_rule <- NULL
for (n_tree in n_tree_values) {
  for (n_split in n_split_values){
    for (split_rule in split_rule_values){
      rfsrc.obj <- rfsrc(surv, data = all_data, nsplit = as.integer(n_split), ntree = n_tree, splitrule = split_rule)
      prederr <- pec(list(rfsrc.obj), data= all_data, splitMethod = "cv10", B = 20, reference = FALSE,formula = surv)$AppErr$rfsrc
      for (i in prederr) {
          if (!is.nan(i)) {
            score<-i
          }
      }
      print(score)
      if (is.null(x) || x>score ) {
        x<-score
        best_n_tree <-n_tree
        best_n_split <- n_split
        best_split_rule <- split_rule
      }
    }
  }
}

res <- list(x,best_n_tree,best_n_split,best_split_rule)
print(res)
```


## Entrainement du modéle avec les hyperparamétres selectionés avec la cross-validation precedentes

```{r message=FALSE, warning=FALSE}
v.out <- rfsrc(surv,all_data,nsplit = res[[3]], ntree=res[[2]], forest=T, splitrule = res [[4]],tree.err=TRUE,importance=TRUE)
print(v.out)
prederr<-pec(list(v.out), data= all_data, splitMethod = "cv10", B = 20, reference = FALSE,maxtime = 400)
```

## Affichage de la coube du score de Brier
```{r}
plot(prederr)
```

```{r}
plot.survival.rfsrc(v.out,plots.one.page = FALSE)
```

```{r}
#ggRandomForest
library("ggplot2")
library("RColorBrewer")
library("dplyr")
library("ggRandomForests")


gg_dta <- gg_survival(interval = "OS",censor = "STATUS",data = all_data,conf.int = 0.95)
plot(gg_dta) + labs(y = "Survival Probability", x = "Observation Time (month)") + theme(legend.position = c(0.2, 0.2)) + coord_cartesian(y = c(0, 1.01))

gg_dta_PS <- gg_survival(interval = "OS",censor = "STATUS",by="PS",data = all_data,conf.int = 0.95)
plot(gg_dta_PS) + labs(y = "Survival Probability", x = "Observation Time (month)",color = "PS", fill = "PS") + theme(legend.position = c(0.2, 0.2)) + coord_cartesian(y = c(0, 1.01))

gg_dta_RMH <- gg_survival(interval = "OS",censor = "STATUS",by="RMH",data = all_data,conf.int = 0.95)
plot(gg_dta_RMH) + labs(y = "Survival Probability", x = "Observation Time (month)",color = "RMH", fill = "RMH") + theme(legend.position = c(0.2, 0.2)) + coord_cartesian(y = c(0, 1.01))
```


```{r}

plot(gg_dta, type = "cum_haz") + labs(y = "Cumulative Hazard", x = "Observation Time (month)") + theme(legend.position = c(0.2, 0.8)) + coord_cartesian(ylim = c(-0.02, 1.22))

plot(gg_dta_PS, type = "cum_haz") + labs(y = "Cumulative Hazard", x = "Observation Time (month)", color = "PS", fill = "PS") + theme(legend.position = c(0.2, 0.8)) + coord_cartesian(ylim = c(-0.02, 1.22))

plot(gg_dta_RMH, type = "cum_haz") + labs(y = "Cumulative Hazard", x = "Observation Time (month)", color = "RMH", fill = "RMH") + theme(legend.position = c(0.2, 0.8)) + coord_cartesian(ylim = c(-0.02, 1.22))
```

```{r}
strCol <- brewer.pal(3, "Set1")[c(2,1,3)]
ggRFsrc <- plot(gg_rfsrc(v.out), alpha = 0.2) +
 scale_color_manual(values = strCol) +
 theme(legend.position = "none") +
 labs(y = "Survival Probability", x = "Time (month)") +
 coord_cartesian(ylim = c(-0.01, 1.01))

 show(ggRFsrc)

```
Random forest OOB predicted survival. Blue curves correspond to censored observations, red curves correspond to observations experiencing death events.
```{r}
plot(gg_rfsrc(v.out, by = "PS")) +
theme(legend.position = c(0.2, 0.2)) +
labs(y = "Survival Probability", x = "Time (month)") +
ggtitle("PS") + 
coord_cartesian(ylim = c(-0.01, 1.01))

```

```{r}
plot(gg_vimp(v.out)) + theme(legend.position = c(0.8, 0.2)) + labs(fill = "VIMP > 0")

```
Random forest Variable Importance (VIMP). Blue bars indicates positive VIMP,
red indicates negative VIMP. Importance is relative to positive length of bars.
VIMP for a variable x is the difference between prediction error when x is randomly permuted, compared to
prediction error under the observed values
Since VIMP is the difference in OOB prediction error before and after permutation, a large
VIMP value indicates that misspecification detracts from the predictive accuracy in the forest.
VIMP close to zero indicates the variable contributes nothing to predictive accuracy, and
negative values indicate the predictive accuracy improves when the variable is misspecified.
In the later case, we assume noise is more informative than the true variable. As such, we
ignore variables with negative and near zero values of VIMP, relying on large positive values
to indicate that the predictive power of the forest is dependent on those variables
```{r}
varsel <- var.select(v.out)

```
```{r}
gg_md <- gg_minimal_depth(varsel)

plot(gg_md)

```
Minimal Depth variable selection. Low minimal depth indicates important variables. The dashed line is the threshold of maximum value for variable selection.
ranking variables from most important at the top (minimal depth measure), to least at the bottom
(maximal minimal depth). The vertical dashed line indicates the minimal depth threshold
where smaller minimal depth values indicate higher importance and larger values indicate
lower importance.

```{r}
plot(gg_minimal_vimp(gg_md)) + theme(legend.position=c(0.8, 0.2))
```
Comparing Minimal Depth and Vimp rankings. Points on the red dashed line
are ranked equivalently, points above have higher VIMP ranking, those below have higher
minimal depth ranking.
The points along the red dashed line indicate where the measures are in agreement. Points
above the red dashed line are ranked higher by VIMP than by minimal depth, indicating the 
variables are more sensitive to misspecification. Those below the line have a higher minimal
depth ranking, indicating they are better at dividing large portions of the population. The
further the points are from the line, the more the discrepancy between measures




As random forest is not parsimonious, we have used minimal depth and VIMP to reduce
the number of variables to a manageable subset. Once we have an idea of which variables
contribute most to the predictive accuracy of the forest, we would like to know how the
response depends on these variables
```{r}

ggRFsrc + geom_vline(aes(xintercept = 1), linetype = "dashed") +
geom_vline(aes(xintercept = 10), linetype = "dashed") +
coord_cartesian(xlim = c(0, 15))


```

```{r}
gg_v <- gg_variable(v.out, time = c(1,10), time.labels = c("1 month", "10 month"))

event.labels <- c(FALSE, TRUE)
event.marks <- c(1, 4)

plot(gg_v, xvar = "CD3_MP_AC_DENSMOY", alpha = 0.4) + #, se=FALSE 
labs(y = "Survival", x = "CD3_MP_AC_DENSMOY") +
theme(legend.position = "none") +
scale_color_manual(values = strCol, labels = event.labels) +
scale_shape_manual(values = event.marks, labels = event.labels) +
coord_cartesian(ylim = c(-0.01, 1.01))
```
Variable dependence of survival at 1 and 10 month on Age_biopsie variable. Individual cases
are marked with blue circles (alive or censored) and red x (dead). Loess smooth curve with
shaded 95% confidence band indicates decreasing survival with increasing bilirubin.


```{r}
plot(gg_v, xvar = "PS", alpha = 0.4) + #, se=FALSE 
labs(y = "Survival", x = "PS") +
theme(legend.position = "none") +
scale_color_manual(values = strCol, labels = event.labels) +
scale_shape_manual(values = event.marks, labels = event.labels) +
coord_cartesian(ylim = c(-0.01, 1.01))
```
```{r}
xvar <- c("CD3_MP_AC_DENSMOY", "CD8_MP_AC_DENSMOY","FOXP3_MP_AC_DENSMOY")

plot(gg_v, xvar = xvar, panel = TRUE, alpha = 0.4) +
labs(y = "Survival") +
theme(legend.position = "none") +
scale_color_manual(values = strCol, labels = event.labels) +
scale_shape_manual(values = event.marks, labels = event.labels) +
coord_cartesian(ylim = c(-0.05, 1.2))

```

```{r}
#partial dependence
time_index <- c(which(v.out$time.interest > 1)[1]-1,
which(v.out$time.interest > 10)[1]-1,
which(v.out$time.interest > 15)[1]-1)


xvar <- c("CD3_MP_AC_DENSMOY", "CD8_MP_AC_DENSMOY","FOXP3_MP_AC_DENSMOY","PS")
partial_pbc <- mclapply(v.out$time.interest[time_index],
function(tm){
plot.variable(v.out, surv.type = "surv",
time = tm, xvar.names = xvar,
partial = TRUE ,
show.plots = FALSE)
 })

gg_dta <- mclapply(partial_pbc, gg_partial)

pbc_ggpart <- combine.gg_partial(gg_dta[[1]], gg_dta[[2]], lbls = c("1 month", "10 month"))

ggplot(pbc_ggpart[["PS"]], aes(y=yhat, x=PS, col=group)) +
geom_boxplot(notch = TRUE, outlier.shape = NA) +
labs(x = "PS", y = "Survival (%)", color="Time", shape="Time") +
theme()

```

```{r}
ggplot(pbc_ggpart[["CD3_MP_AC_DENSMOY"]], aes(y=yhat, x=CD3_MP_AC_DENSMOY, col=group)) +
geom_point() + geom_smooth() +
labs(x = "CD3_MP_AC_DENSMOY", y = "Survival (%)", color="Time", shape="Time") +
theme()

```

```{r}
#variable interactions
ggint <- gg_interaction(v.out)
xvar <- c("CD3_MP_AC_DENSMOY", "CD8_MP_AC_DENSMOY","FOXP3_MP_AC_DENSMOY","PS")
plot(ggint, xvar = xvar)

```

```{r}
#conditional dependence plots
ggvar <- gg_variable(v.out, time = 10)
ggvar$PS <- paste("PS = ", ggvar$PS, sep = "")

var_dep <- plot(ggvar, xvar = "CD3_MP_AC_DENSMOY",
alpha = 0.5) +
# geom_smooth(method = "glm",se = FALSE) +
labs(y = "Survival at 10 month",
x = "CD3_MP_AC_DENSMOY") +
theme(legend.position = "none") +
scale_color_manual(values = strCol, labels = event.labels) +
scale_shape_manual(values = event.marks, labels = event.labels) +
coord_cartesian(y = c(-.01,1.01))

var_dep
```
```{r}
var_dep + facet_grid(~PS)
```

